PRISM protocol v0.3

### 2021-09-23 prism-backend/research/article.tex
ATA-4755: add section addressing scalability (#150)
This commit adds a section exploring scalability proposals for our
protocol

% \section{Ideas to add missing properties}

\end{document}

>>>>

% \section{Ideas to add missing properties}
\section{Further thoughts}

\subsection{Scaling}

In the previous sections we described a protocol that allows transparent management of decentralized identifiers and credentials.
The protocol usage of a blockchain facilitates the solution to three points:
\begin{itemize}
\item Events ordering: All participants can agree on the order in which any pair of events occurred 
\item Fork protection (safety): There is no point in time where a valid sequence of processed protocol events could become invalid, nor replaced by 
      another sequence
\item Data transmission: All participants receive the same sequence of protocol events as they are transmitted through the blockchain
\end{itemize}

The second point has particular relevance to guarantee that the history of events can be trusted. For instance, imagine Alice controls a DID (its 
associated master keys) that represents a real world object, like a car. If Alice's wants to sell the car and transfer the corresponding DID, she 
would perform a DID update that revokes all the master keys controlled by her, and adds a new master key controlled by the new car owner. If Alice 
could later "undo" this update, the DID ownership transfer could not be trusted.

The properties brought by the blockchain come at a cost. Namely, the protocol throughput (number of events that participants can attach in 
transactions per unit of time) is bounded to the blockchain throughput. In order to scale the throughput, we have reviewed Sidetree's approach. The 
idea behind Sidetree is to not publish individual events attached to blockchain transactions, but to publish a hash-link to an off-chain content 
addressable storage (CAS), like IPFS. The CAS will contain the file referenced by the hash-link, which contains a batch of many protocol events 
(allowing to transcend the size limits imposed by blockchain transactions).

We identify a non-ideal drawback about this approach, which is a loss of the safety property. In the setting defined by Sidetree, participants can 
control changes to the past of their identifiers. This is, if Alice desires, she could create a DID by posting a file, $F_{1}$ on the CAS, and its 
hash on-chain, $F_{1}$ would contain Alice $CreateDID$ event. Later, Alice can create a file $F_{2}$, containing an $UpdateDID$ event, post the hash 
of $F_{2}$ on-chain, but intentionally not posting $F_{2}$ on the CAS. Then, Alice can create a third file $F_{3}$ which contains an new $UpdateDID$
event that would be invalid if $F_{2}$ were revealed in the CAS (e.g. $F_{2}$ event could revoke the key that signs the update in $F_{3}$). However, 
Alice can post both $F_{3}$ in the CAS and the hash of $F_{3}$ on-chain. In Sidetree, this sequence of actions would lead all protocol participants 
to believe that Alice's DID is in certain state, produced by the $CreateDID$ of $F_{1}$ and the $UpdateDID$ from $F_{3}$.

Alice has the power to reveal at any point $F_{2}$, making the update from $F_{3}$ as invalid, and forcing all participants to update Alice's
DID to the state reflected by $F_{2}$. This is known as \textbf{late publish attack}, and makes technically impossible to trust the past history 
history of events in Sidetree implementations. As an example consequence, it is not possible to transfer ownership of a DID when using Sidetree.

In order to avoid introducing this issue, we considered two options.

The first one is to add a permissioned actor (or a federation of them) that are allowed to batch protocol events and publish both on-chain hash-links,
and files in corresponding CAS. The special actor introduces a trust model that assumes that it will always reveal files. If the actor fails to 
reveal a file, the protocol participants stop processing further batches until the missing file is revealed. This is not possible in Sidetree because
any participant would be able to freeze the system by not revealing a file. The protocol would still allow users to publish events on-chain directly, 
leaving space for some decentralization for those participants that do not want to depend on the centralized batcher.

In order to use the batcher to publish the $UpdateDID$ events associated to a DID $D$, the owner of $D$ will have to declare on-chain that future 
updates for $D$ will be found on the off-chain batches. This public declaration is needed to avoid race conditions when a hash is on-chain but
a file is not revealed.

A second option consist of a more decentralized variation of the previous approach. Namely, allow any participant to propose itself as an event 
batcher. The proposal is performed by submitting an on-chain event declaring the batcher DID. Every user that would like to batch its events 
off-chain should publicly notify the batcher they will use. Only one batcher can be assigned to each DID, this is, the $UpdateDID$ events associated 
to a DID will be published by at most one batcher. The objective is once again, to be able to identify missing files in a sequence, allowing 
participants to stop processing files out of order.

In any of the approaches, if the associated batcher stops publishing files, the DID registered to the batcher is "frozen" until the file is revealed.
We could re-introduce some liveness protection to DID owners by requesting batches from a same batcher to be separated at least $N$ underlying 
blockchain blocks, in order to provide that time for users to send a $Contention$ event that invalidates the events associated to a DID published in
the previous batch. During the processing of batches, participants will wait the contention period (the $N$ blocks) before applying events from a 
batch in order to not apply the updates to contended DIDs.

Notes:
\begin{itemize}
\item we could change the $CreateDID$ event to incorporate an optional batcher DID from start, the batcher DID would be part of the initial state, 
      bounding the batcher to the DID suffix. 
\item we could support events to de-register from a batcher, and to switch batcher too.
\item a user is free to batch his own DIDs' related events. This could be useful for a case like a car manufacturer that would like to batch all DIDs
      associated to their cars, allowing car owners to de-register from a batcher at will.
\item a priori, we do not see a need to batch credential issuance/revocation events
\end{itemize}
\end{document}
